---
title: "Basic inferential data analysis"
author: "Zanin Pavel"
date: "February 28, 2016"
output: pdf_document
---

[Link to project on GitHUB](https://github.com/Filareth2015/Statistical-Inference-Course-Project.git)

# Overview
In this project we are going to analyze the ToothGrowth data from the R datasets package.  
This dataset describes how changes the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

# 1. Getting data and exploratory analysis

For start working load the dataset:
```{r}
library(datasets) #loading neccesary library
data(ToothGrowth) #loading specified dataset
```

Let's see what are these dataset:
```{r}
str(ToothGrowth) #compactly displaying the internal structure 
head(ToothGrowth) #showing the first 6 rows of dataset
summary(ToothGrowth) #showing dataset's summary  
ToothGrowth$dose # showing the list of doses
```

So, as we can see, we've got 60 observations, for 2 supplement types (VC or OJ) and 3 dose levels of Vitamin C (0.5, 1, and 2). Dataset's description not lied to us =)

For the next step let's make exploratory plot's for this data:  
```{r}
library(ggplot2) #loading neccesary library

ggplot(ToothGrowth, aes(x=factor(dose), y=len)) + 
  facet_grid(.~supp) +
  geom_boxplot(aes(fill = supp)) +
  labs(title="Guinea pig tooth length by supplement type 
  (orange juice (OJ) or ascorbic acid (VC))", 
    x="Dose (mg)",
    y="Tooth Length")
```

# 2. Hypothesis testing

## Hypotesis 1: There is no difference between supplement types (orange juice or ascorbic acid), regardless from doses  

For testing this let`s try t.test:
```{r}
t.test(len ~ supp, data = ToothGrowth)
```

So, if confidence interval include zero and p-value is bigger than usual $\alpha$ level (.05) then we our hypotesis is true and we cannot reject it.

## Hypotesis 2: Effect from 0.5 mg dose for both supplement types is equal

For testing this try t.test again:
```{r}
t.test(len ~ supp, data = subset(ToothGrowth, dose == 0.5))
```

As p-value is lower than usual $\alpha$ level (.05) then our hypotesis isn't true and we reject it. Orange juice has much effectiveness for this dose than ascorbic acid.

## Hypotesis 3: Effect from 1 mg dose for both supplement types is equal

For testing this try t.test again:
```{r}
t.test(len ~ supp, data = subset(ToothGrowth, dose == 1))
```

As p-value is lower than usual $\alpha$ level (.05) then our hypotesis isn't true and we reject it. Orange juice has much effectiveness for this dose than ascorbic acid.

## Hypotesis 4: Effect from 2 mg dose for both supplement types is equal

For testing this try t.test again:
```{r}
t.test(len ~ supp, data = subset(ToothGrowth, dose == 2))
```

So, if confidence interval include zero and p-value is bigger than usual $\alpha$ level (.05) then we our hypotesis is true and we cannot reject it.

# 3. Conclusions 
Dataset ToothGrowth allows to us make next conclusions:   
** Vitamin C consumption results to increasing pig's tooth growth.   
** In small doses (0.5 and 1 mg) orange juice much effective than ascorbic acid.    
** In big dose (2 mg) both supply types have same effectiveness.

